\bf Interactive Visualization of Hierarchically Structured Data

### 2017-09-23 reviewer_response_ii.tex

I als want to commend the details presented in the video and on the GitHub
pages. Very nicely done.

I have a few minor suggestions for a few additional revisions that might help
the paper.

1. Regarding the third and fourth view, your response to earlier comments
suggests that implementation details interfered with attempts to combine these
views. Discussing these concerns in the paper (or on the documentation website)
might help others either avoid such difficulties or identify appropriate
work-arounds. Please consider adding relevant text to the paper.

\color{black}
We have added a brief discussion of the possibility of combining the third and
fourth views in a paragraph following their introduction (``In principle, it
would be desirable...''). We describe the difficulties we found, and provide a
link to a github branch with the partial solution that we found unsatisfactory.

@@

\color{blue}
2. On page 7, the description of the tree reads: "Every edge in the tree is
split into two colors1, with relative widths of the different colors reflecting
differences in sizes for the two groups. The overall width of each edge
represents the sum of values across all groups." I was a bit confused by these
sentences, as they seem redundant. Doesn't the second sentence follow directly
from the first? If not, some additional clarification is needed.

\color{black}
It's true, those sentences are redundant. We realize now that this can be
confusing, so we have removed the second sentence.

@@

\color{blue}
3. Regarding the tree ordering, I found myself wondering about the reordering of
the trees based on the size of the subtree. As this seems to be to be a tradeoff
between information clarity and consistency, I can imagine some cases where this
behavior might be useful, and others where it might be distracting. Have you
considered the possibility of animating the transition, so was to help users
avoid disorientation? Alternatively, a toggle to enable/disable this behavior
might help.

\color{black}
We had not previously considered using orderings other than D3's default.
However, we were able to incorporate the reviewer's suggestion by adding a
``toggle layout'' button, which smoothly transitions between D3's default ordering
(descending height and then ascending ID) and an ordering based on the number of
total descendants. The online vignettes have been updated to include this new
functionality.

### 2017-09-23 treelapse.tex
remove supplemental material

We have implemented our visualizations in D3, but encapsulated in an R package,
called treelapse, to facilitate rapid turnover from data preparation and
modeling to interactive exploration, and vice versa. Our code is open-source,
and is linked in Section \ref{sec:supp_links} of the supplementary materials. We
hope this package encourages data analysts to work at the border between data
modeling and visualization, and more generally empowers a wider audience to
apply less widely known, but powerful, visualization ideas.

@@

In the DOI sankey, we split each edge in the DOI Tree across different groups.
For example, suppose we have the average counts for treatment and control groups
at each tree node. Every edge in the tree is split into two colors\footnote{We
  use the colorbrewer palette to facilitate readability
  \citep{brewer2003colorbrewer}}, with relative widths of the different colors
reflecting differences in sizes for the two groups. The overall width of each
edge represents the sum of values across all groups.

@@

In principle, it would be desirable to combine the timebox and treebox displays,
so that highlighted nodes and series could be determined through brushes on both
the tree and series. For example, it would be useful to highlight the samples
that lie in the intersection of all timeboxes and union of all treeboxes.
However, the authors have found practical implementation of this idea
challenging.

@@

We next consider an application unrelated to the microbiome, but with relatively
clear hierarchical structure. Our data are downloaded from Zillow and give the
Zillow Home Value Indexes at the neighborhood level, across the country,
computed monthly between 1996 and 2016. A link to the data source is provided in
Section \ref{sec:supp_links} of the supplementary materials. In our display, we
have taken the natural log of these indexes. As our hierarchical structure, we
use each neighborhood's assignment to state, regional, county, and city levels.
We represent each of these coarser spatial categories using the average of all
neighborhoods contained in them. We have filtered down to the 890 neighborhoods
in California; rendering more neighborhoods while keeping all 246 timepoints
causes a severe lag in the interface.

@@

Our next example is a study in bikesharing demand, included as an example of
analyzing collections of time series when there is no obvious hierarchical
structure a priori. The data are available at the UCI Repository and were
originally collected by a Washington D.C.-based bikesharing system for use in a
Kaggle prediction competition. A link is provided in Section
\ref{sec:supp_links} of the supplementary materials. The data are hourly
measurements of bike demand, aggregated across all bikesharing stations, over
two years, along with supplemental weather data. In the competition,
participants were asked to predict the hourly demand on a held-out test set.
Here, we adopt a descriptive view instead, attempting to characterize factors
associated with variation in bikesharing demand.

@@

\section{Supplemental Materials}

\subsection{Links}
\label{sec:supp_links}

\begin{itemize}
\item treelapse home page: \url{https://krisrs1128.github.io/treelapse/}
\item treelapse development page: \url{https://github.com/krisrs1128/treelapse}
\item treelapse video demo: \url{https://www.youtube.com/watch?v=EcmYBRMVMbI&t=14s}
\item Home prices data: \url{http://www.zillow.com/research/data/}
\item Bikesharing data: \url{https://archive.ics.uci.edu/ml/datasets/Bike+Sharing+Dataset}
\end{itemize}

>>>>

We have implemented our visualizations in D3, but encapsulated in an R package,
called treelapse, to facilitate rapid turnover from data preparation and
modeling to interactive exploration, and vice versa. Our code is open-source,
and is linked in the supplementary materials. We hope this package encourages
data analysts to work at the border between data modeling and visualization, and
more generally empowers a wider audience to apply less widely known, but
powerful, visualization ideas.

@@

In the DOI sankey, we split each edge in the DOI Tree across different groups.
For example, suppose we have the average counts for treatment and control groups
at each tree node. Every edge in the tree is split into two colors\footnote{We
  use the colorbrewer palette to facilitate readability
  \citep{brewer2003colorbrewer}}, with relative widths of the different colors
reflecting differences in sizes for the two groups.

@@

In principle, it would be desirable to combine the timebox and treebox displays,
so that highlighted nodes and series could be determined through brushes on both
the tree and series. For example, it would be useful to highlight the samples
that lie in the intersection of all timeboxes and union of all treeboxes. This
could allow more complex queries than are currently available. However, while
conceptually appealing, the authors encountered obstacles in practical
implementation: brush and mouseover events are required to occupy the same space
in this combined view. Properly distinguishing these events can be challenging,
and a solution based on the introduction of a lag between mousedown and the
manipulation of a brush led to a deteriorated user experience\footnote{However,
  the code for this approach is available publicly, in a separate branch of
  treelapse:
  \url{https://github.com/krisrs1128/treelapse/tree/combined-brushes}.}.

@@

We next consider an application unrelated to the microbiome, but with relatively
clear hierarchical structure. Our data are downloaded from Zillow and give the
Zillow Home Value Indexes at the neighborhood level, across the country,
computed monthly between 1996 and 2016. A link to the data source is provided in
the supplementary materials. In our display, we have taken the natural log of
these indexes. As our hierarchical structure, we use each neighborhood's
assignment to state, regional, county, and city levels. We represent each of
these coarser spatial categories using the average of all neighborhoods
contained in them. We have filtered down to the 890 neighborhoods in California;
rendering more neighborhoods while keeping all 246 timepoints causes a severe
lag in the interface.

@@

Our next example is a study in bikesharing demand, included as an example of
analyzing collections of time series when there is no obvious hierarchical
structure a priori. The data are available at the UCI Repository and were
originally collected by a Washington D.C.-based bikesharing system for use in a
Kaggle prediction competition. A link is provided in the supplementary
materials. The data are hourly measurements of bike demand, aggregated across
all bikesharing stations, over two years, along with supplemental weather data.
In the competition, participants were asked to predict the hourly demand on a
held-out test set. Here, we adopt a descriptive view instead, attempting to
characterize factors associated with variation in bikesharing demand.

@@
